
# Time Series Forecasting
# Kamal | Sep-2016

### 1: Load required libraries

```{r setup}
library(dplyr)
library(tseries)
```

### 2: Loading the Data Set

```{r}

data(AirPassengers)
class(AirPassengers) #This tells you that the data series is in a time series format
start(AirPassengers) #This is the start of the time series
end(AirPassengers) #This is the end of the time series
frequency(AirPassengers) #The cycle of this time series is 12months in a year
summary(AirPassengers)

```

### 3: Exploratory Metrics Analysis

```{r}
plot(AirPassengers)
abline(reg=lm(AirPassengers~time(AirPassengers))) # This will fit in a line
cycle(AirPassengers) #This will print the cycle across years.
plot(aggregate(AirPassengers,FUN=mean)) #This will aggregate the cycles and display a year on year trend
boxplot(AirPassengers~cycle(AirPassengers)) #Box plot across months will give us a sense on seasonal effect
```

* So key inferences that we can get are as follows:
*  The year on year trend clearly shows that the no of passengers have been increasing without fail.
*  The variance and the mean value in July and August is much higher than rest of the months.
*  Even though the mean value of each month is quite different their variance is small. Hence, we have strong seasonal effect with a cycle of 12 months or less.

* Above exploratory data analysis is important as it can indicate whether a series is stationary ornot.

* Try with Augmented Dickey-Fuller Test
* We know that we need to address two issues before we test stationary series. One, we need to remove unequal variances. We do this using log of the series. 
* Two, we need to address the trend component. We do this by taking difference of the series. Now, let’s test the resultant series.
* We see that the series is stationary enough to do any kind of time series modelling.

```{r}
adf.test(diff(log(AirPassengers)), alternative="stationary", k=0)

```

### 4: Find Optimal parameters

```{r}
# ACF plots to be drawn
acf(log(AirPassengers))
# Clearly, the decay of ACF chart is very slow, which means that the population is not stationary. 
# We now intend to regress on the difference of logs rather than log directly. Let’s see how ACF and PACF 
# curve come out after regressing on the difference.
acf(diff(log(AirPassengers)))
pacf(diff(log(AirPassengers)))

```

### 5: Build and Fit ARIMA model

```{r}
# Let’s fit an ARIMA model and predict the future 10 years. Also, we will try fitting in a seasonal component 
# in the ARIMA formulation. Then, we will visualize the prediction along with the training data. 
(fit <- arima(log(AirPassengers), c(0, 1, 1),seasonal = list(order = c(0, 1, 1), period = 12)))
```

### 6: Make Predictions

```{r}
pred <- predict(fit, n.ahead = 10*12)
ts.plot(AirPassengers,2.718^pred$pred, log = "y", lty = c(1,3))
```
